\textbf{Proof Reconstruction in Classical Propositional Logic}

### 2017-04-28 slides/slides.tex
[ slides ] example proof added.

% --------------------------------------------------------------------------
% Packages
% --------------------------------------------------------------------------
\usepackage[english]{babel}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{bussproofs}
\EnableBpAbbreviations
\usepackage{fancyvrb}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{lmodern}
\usepackage{mylst}
\usepackage{url}

@@

\usepackage[autostyle]{csquotes}
\usepackage[
    backend=biber
  , style=authoryear-icomp
  , sortlocale=en_US
  , natbib=true
  , url=false
  , doi=true
  , eprint=false
]{biblatex}
\addbibresource{ref.bib}
\renewcommand*{\nameyeardelim}{\addcomma\addspace}
\usepackage{silence}
\WarningFilter{biblatex}{Patching footnotes failed}
\WarningFilter{hyperref}{Token not allowed in a PDF string}

@@

\begin{frame}[fragile, label=reconstruction]{Proof Reconstruction: Overview}
\vfill
\begin{tikzpicture}
\node(problem) at (0,0)
   {\includegraphics[scale=0.7]{figures/problem}};
\node[right = 0.8cm of problem, hyperlink node=tptp-syntax](tptp){
\includegraphics[scale=0.7]{figures/tptp}
};
\node[right= 0.8cm of tptp, hyperlink node=metis] (metis)
  {\includegraphics[scale=0.7]{figures/metis}\,
  \footnote{It is available at \url{http://www.gilith.com/software/metis}}
  };
\node[right= 0.8cm of metis, hyperlink node=tstp-syntax] (tstp)
  {\includegraphics[scale=0.7]{figures/tstp}};
\node[below= 1.5cm of tstp, hyperlink node=athena] (athena)
  {\includegraphics[scale=0.7]{figures/athena}
  \footnote{It is available at \url{http://github.com/jonaprieto/athena}}
  };
\node[left = 0.5cm of athena, hyperlink node=agda-file] (agdafile)
  {\includegraphics[scale=0.7]{figures/agdafile}
  };
\node[left = 0.5cm of agdafile, hyperlink node=agda] (agda)
  {\includegraphics[scale=0.7]{figures/agda}
  \footnote{It is available at \url{http://github.com/agda/agda}}
  };
\node[below = 0.18cm of problem, hyperlink node=verified-example] (verified)
  {\includegraphics[scale=0.7]{figures/verified}};
\node[below = 0.18cm of verified, opacity=0.5, hyperlink node=failure-example] (failure)
  {\includegraphics[scale=0.7]{figures/failure}};

@@

@@

A TSTP derivation
\footnote{\url{http://www.cs.miami.edu/~tptp/TPTP/QuickGuide/Derivations.html}}
\begin{itemize}
  \item Is a \hyperlink{tstp-dag}{\textbf{D}irected \textbf{A}cyclic \textbf{G}raph} where
  \begin{itemize}
    \item[\texttt{leaf}] is a formula from the \hyperlink{tptp-syntax}{TPTP} input
    \item[\texttt{node}] is a formula inferred from parent formula
    \item[\texttt{root}] the final derived formula
  \end{itemize}
  \item Is a list of annotated formulas with the form:
  \end{itemize}

@@

where \textbf{source} typically is an inference record:
\begin{center}
{\footnotesize
\begin{tptp}
inference(rule, useful info, parents)
\end{tptp}
}
\end{center}
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=tstp-example]{TSTP Example}

@@

\begin{itemize}
  \item Proof found by \texttt{Metis} ATP for the problem $p ⊢ p$
{\small
\begin{tptp}
$ metis --show proof basic-4.tptp
fof(a, axiom, (p)).
fof(goal, conjecture, (p)).
fof(subgoal_0, plain, (p),
  inference(strip, [], [goal])).
fof(negate_0_0, plain, (~ p),
  inference(negate, [], [subgoal_0])).
fof(normalize_0_0, plain, (~ p),
  inference(canonicalize, [], [negate_0_0])).
fof(normalize_0_1, plain, (p),
  inference(canonicalize, [], [a])).
fof(normalize_0_2, plain, ($false),
  inference(simplify, [],
    [normalize_0_0, normalize_0_1])).
cnf(refute_0_0, plain, ($false),
    inference(canonicalize, [], [normalize_0_2])).
\end{tptp}
}
\end{itemize}
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=tstp-dag]{DAG for the previous TSTP derivation found by \texttt{Metis} ATP}

\vfill
\begin{center}
\includegraphics[height=0.8\textheight]{figures/derivation.png}
\end{center}
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=athena]{Athena}
 \tikz[overlay,remember picture]
  \node at (0.92\textwidth, 1cm)
    {\includegraphics[width=0.15\textwidth]{figures/athena}};
\begin{block}{Athena} Is a Haskell program that translates proofs given by Metis Prover
in TSTP format to Agda code.\\
It depends on:
\begin{itemize}
  \item \texttt{agda-prop} Classical Logic within Agda: Axioms + Theorems
  \item \texttt{agda-metis} Theorems of the inference rules of Metis Prover
\end{itemize}
\end{block}
\gobackreconstruction
\end{frame}

\begin{frame}[label=haskell]{Design Decisions for the Reconstruction Tool}
 \tikz[overlay,remember picture]
  \node at (0.93\textwidth, 0cm)
    {\includegraphics[width=0.2\textwidth]{figures/athena}};

\begin{block}{Haskell} % is a standardized, general-purpose purely functional programming language.\\ Our usages:
\begin{itemize}
    \item Parsing
    \item AST construction
    \item Creation and analysis of \hyperlink{tstp-dag}{DAG} derivations
    \item Analysis of inference rules used
    \item Generation of Agda code of the proof
\end{itemize}
\end{block}
\gobackreconstruction
\end{frame}

\begin{frame}[label=agda]{Design Decisions for the Reconstruction Tool}{Programming Languages}

 \tikz[overlay,remember picture]
  \node at (0.92\textwidth, 1cm)
    {\includegraphics[width=0.15\textwidth]{figures/agda}};

\begin{block}{Agda}
Is a dependently typed functional programming language and it also a proof assistant.\\
We used it to type-check the proofs found by Metis Prover
    \begin{itemize}
    \item Agda-Prop Libary: Logic framework for Classical Propositional Logic
    \item Agda-Metis Library: theorems based on the inference rules of Metis Prover
    \end{itemize}
\end{block}
\gobackreconstruction
\end{frame}

Metis is an automatic theorem prover for First-Order Logic with equality
\begin{block}{Why Metis?}
  \begin{itemize}
    \item \href{https://github.com/gilith/metis}{Open source} implemented in Standard ML
    \item Reads problem in \hyperlink{tptp-syntax}{\texttt{TPTP} format}
    \item Outputs detailed proofs in \hyperlink{tstp-format}{\texttt{TSTP} format}
    \item Each refutation step is one of
      \href{http://www.gilith.com/software/metis/logical-kernel.txt}{\textbf{6 simple rules}}
  \end{itemize}
\end{block}
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=metis-rules]{Inference Rules of Metis}
\hyperlink{tstp-example}{TSTP} derivations exhibit these inferences:
\vfill
\begin{table}[!ht]
\begin{center}
\scalebox{0.8}{
{\renewcommand{\arraystretch}{1.3}%
\begin{tabular}{|cl| }
\hline
\textbf{Rule} & \textbf{Task} \\
\hline
\texttt{canonicalize} & transforms formulas to CNF, DNF or NNF\\
\texttt{clausify} & performs clausification\\
\hyperlink{atp-conjunct}{\texttt{conjunct}} & extracts a formula from a conjunction\\
\texttt{negate} & applies negation to the formula\\
\texttt{resolve} & applies theorems of resolution\\
\texttt{simplify} & applies over a list of formula to simplify them\\
\texttt{strip} & splits a formula into subgoals\\[1ex]
\hline
\end{tabular}}
}
\end{center}
\label{tab:label}
\end{table}
\vfill
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=atp-conjunct]{Conjunct}{Metis Inference Rule}

\begin{itemize}
\item Definition
  \begin{equation*}
  conjunct (ϕ₁ ∧  ϕ₂ ∧ ⋯ ∧ \mathbf{ϕᵢ} ∧ ⋯ ∧ ϕₙ, \mathbf{ϕᵢ}) ⟶ \mathbf{ϕᵢ}
  \end{equation*}

\item Function\footnote{\label{note-conjunct}It is available in \texttt{agda-metis} library in \texttt{ATP.Metis.Rules.Conjunct} module.}:
\begin{agda}
conjunct : Prop → Prop → Prop
conjunct $(φ \wedge ψ)$ $ω$ with $⌊$ eq $φ$ $ω$ $⌋$ | $⌊$ eq ψ $ω$ $⌋$
... | true  | _     = $φ$
... | false | true  = $ψ$
... | false | false = conjunct $φ$ $ω$
conjunct $φ$ $ω$       = $φ$
\end{agda}
\item Theorem\textsuperscript{\ref{note-conjunct}}
\begin{agda}
atp-conjunct
  : ∀ $\{Γ\}$ $\{φ\}$
  → $(ω : Prop)$
  → $Γ \vdash φ$
  → $Γ \vdash$ conjunct $φ$ $ω$
\end{agda}
\end{itemize}
\end{frame}

@@

\begin{itemize}
  \item The problem is $p ∧ q ⊢ q ∧ p$
  \item TPTP derivation
\begin{tptp}
fof(a, axiom, p & q).
fof(goal, conjecture, q & p).
\end{tptp}
  \item A natural deduction proof
\begin{columns}
\begin{column}{0.8\textwidth}
 \begin{prooftree}
  \AxiomC{$ϕ ∧ ψ$}
  \RightLabel{$∧$-\texttt{proj}$₁$}
  \UnaryInfC{$ϕ$}

@@

% \begin{frame}[fragile, label=verified-example1]{TPTP}

% \end{frame}

\begin{frame}[fragile, label=verified-example]{Type-checked Proof}
\vfill
\begin{agda}
proof$₀$ : $Γ ⊢$ subgoal$₀$
proof$₀$ =
  (RAA
    (atp-canonicalize
      (atp-simplify
        (atp-canonicalize
          (atp-strip
            (assume $\{Γ = Γ\}$
              (atp-negate subgoal$₀$))))
        (atp-conjunct (q)
          (atp-canonicalize
            (weaken (atp-negate subgoal$₀$)
              (assume $\{Γ = ∅\}$ a)))))))
\end{agda}
\vfill
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=verified-example-2]{Type-checked Proof}
\vfill
\begin{agda}
proof$₁$ : $Γ ⊢$ subgoal$₁$
proof$₁$ =
  (RAA
    (atp-canonicalize
      (atp-simplify
        (atp-conjunct (q)
          (atp-canonicalize
            (weaken (atp-negate subgoal$₁$)
              (assume $\{Γ = ∅\}$ a))))
        (atp-simplify
          (atp-canonicalize
            (atp-strip
              (assume $\{Γ = Γ\}$
                (atp-negate subgoal$₁$))))
          (atp-conjunct (p)
            (atp-canonicalize
              (weaken (atp-negate subgoal$₁$)
                (assume $\{Γ = ∅\}$ a))))))))
\end{agda}
\vfill
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=failure-example]{Type-checked Proof}
Failure Example goes here
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=hammer]{Related Work}
\begin{block}{SledgeHammer}
\begin{itemize}
\item Isabelle/HOL tool
\item Metis ported within Isabelle
\item Reconstruct proofs of well-known ATPs: EProver, Vampire, among others
\end{itemize}
\end{block}

\begin{block}{Integrating Waldmeister and Agda}
\begin{itemize}
  \item Source code not available
  \item Equational Logic
  \item Reflection Layers
  \end{itemize}
\end{block}
\end{frame}

@@

\begin{frame}[fragile, label=after-reconstruction]
\end{frame}

@@

\only<4->{
  \node[left= 1cm of apia] (tptp)
    {\includegraphics[scale=0.8]{figures/tptp}}};

@@

\only<7->{ \node[below = 2.4 of agdafile, rectangle, text=white,fill=black, align=left, font=\tiny] (proved)
{
\texttt{\$ agda Or.agda}\\
\texttt{\$ apia Or.agda --atp=online-metis}\\
\texttt{Metis---2.3 proved the conjecture}
  }
};

@@

\begin{frame}[label=future-work]{Future Work}
\begin{itemize}
  \item Add shallow embedding in order to work with \texttt{Apia}
  \item Support First-Order Logic with Equality
  \item Support another prover like \texttt{EProver}\textbf{
}\end{itemize}
\end{frame}

>>>>

% --------------------------------------------------------------------------
% Packages
% --------------------------------------------------------------------------
\usepackage[english]{babel}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{bussproofs}
\EnableBpAbbreviations
\def\extraVskip{3pt}

\usepackage{fancyvrb}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{lmodern}
\usepackage{mylst}
\usepackage{url}

@@

\usepackage[autostyle]{csquotes}
\usepackage[
    backend=biber
  , style=authoryear-icomp
  , sortlocale=en_US
  , natbib=true
  , url=false
  , doi=false
  , isbn=false
  , eprint=false
]{biblatex}
\addbibresource{ref.bib}
\renewcommand*{\nameyeardelim}{\addcomma\addspace}
\usepackage{silence}
\WarningFilter{biblatex}{Patching footnotes failed}
\WarningFilter{hyperref}{Token not allowed in a PDF string}

@@

\begin{frame}[fragile, label=reconstruction]{Proof Reconstruction: Overview}
\vfill
\begin{tikzpicture}
\node(problem) at (0,0)
   {\includegraphics[scale=0.7]{figures/problem}};
\node[right = 0.8cm of problem, hyperlink node=tptp-syntax](tptp){
\includegraphics[scale=0.7]{figures/tptp}
};
\node[right= 0.8cm of tptp, hyperlink node=metis] (metis)
  {\includegraphics[scale=0.7]{figures/metis}};
\node[right= 0.8cm of metis, hyperlink node=tstp-syntax] (tstp)
  {\includegraphics[scale=0.7]{figures/tstp}};
\node[below= 1.5cm of tstp, hyperlink node=athena] (athena)
  {\includegraphics[scale=0.7]{figures/athena}};
\node[left = 0.5cm of athena, hyperlink node=agda-file] (agdafile)
  {\includegraphics[scale=0.7]{figures/agdafile}};
\node[left = 0.5cm of agdafile, hyperlink node=agda] (agda)
  {\includegraphics[scale=0.7]{figures/agda}};
\node[below = 0.18cm of problem, hyperlink node=verified-example] (verified)
  {\includegraphics[scale=0.7]{figures/verified}};
\node[below = 0.18cm of verified, opacity=0.5, hyperlink node=failure-example] (failure)
  {\includegraphics[scale=0.7]{figures/failure}};

@@

Metis is an automatic theorem prover for First-Order Logic with equality
\begin{block}{Why Metis?}
  \begin{itemize}
    \item \href{https://github.com/gilith/metis}{Open source} implemented in Standard ML
    \item Each refutation step is one of
      \href{http://www.gilith.com/software/metis/logical-kernel.txt}{\textbf{6 simple rules}}
    \item Reads problem in \hyperlink{tptp-syntax}{\texttt{TPTP} format}
    \item Outputs detailed proofs in \hyperlink{tstp-syntax}{\texttt{TSTP} format}
  \end{itemize}
\end{block}
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=metis-rules]{Inference Rules of Metis}
\hyperlink{tstp-example}{TSTP} derivations by Metis exhibit these inferences
\footnote{Inference rules found in proofs of Propositional Logic theorems}:
\vfill
\begin{table}[!ht]
\begin{center}
\scalebox{0.8}{
{\renewcommand{\arraystretch}{1.3}%
\begin{tabular}{|cl| }
\hline
\textbf{Rule} & \textbf{Purpose} \\
\hline
\texttt{canonicalize} & transforms formulas to CNF, DNF or NNF\\
\texttt{clausify} & performs clausification\\
\hyperlink{atp-conjunct}{\texttt{conjunct}} & takes a formula from a conjunction\\
\texttt{negate} & applies negation to the formula\\
\texttt{resolve} & applies theorems of resolution\\
\texttt{simplify} & applies over a list of formula to simplify them\\
\texttt{strip} & splits a formula into subgoals\\[1ex]
\hline
\end{tabular}}
}
\end{center}
\label{tab:label}
\end{table}
\vfill
\gobackreconstruction
\end{frame}

@@

A TSTP derivation
\footnote{\url{http://www.cs.miami.edu/~tptp/TPTP/QuickGuide/Derivations.html}}
\begin{itemize}
  \item Is a \hyperlink{tstp-dag}{\textbf{D}irected \textbf{A}cyclic \textbf{G}raph} where
  \begin{itemize}
    \item[\texttt{leaf}] is a formula from the \hyperlink{tptp-syntax}{\texttt{TPTP}} input
    \item[\texttt{node}] is a formula inferred from parent formula
    \item[\texttt{root}] the final derived formula
  \end{itemize}
  \item Is a list of annotated formulas with the form:
  \end{itemize}

@@

where \textbf{\color{plum} source} typically is an inference record:
\begin{center}
{\footnotesize
\begin{tptp}
inference(rule, useful info, parents)
\end{tptp}
}
\end{center}
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=tstp-example]{\texttt{TSTP} Example}

@@

\begin{itemize}
  \item Proof found by \texttt{Metis} Prover for the problem $p ⊢ p$
{\small
\begin{tptp}
$ metis --show proof basic-4.tptp
fof(a, axiom, p).
fof(goal, conjecture, p).
fof(subgoal_0, plain, p),
  inference(strip, [], [goal])).
fof(negate_0_0, plain, ~ p,
  inference(negate, [], [subgoal_0])).
fof(normalize_0_0, plain, ~ p,
  inference(canonicalize, [], [negate_0_0])).
fof(normalize_0_1, plain, p,
  inference(canonicalize, [], [a])).
fof(normalize_0_2, plain, $false,
  inference(simplify, [],
    [normalize_0_0, normalize_0_1])).
cnf(refute_0_0, plain, $false,
    inference(canonicalize, [], [normalize_0_2])).
\end{tptp}
}
\end{itemize}
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=tstp-dag]{DAG for the previous TSTP derivation found by \texttt{Metis} Prover}

\vfill
By refutation, we proved $p ⊢ p$:
\begin{center}
\begin{prooftree}
\AxiomC{$p$}
\RightLabel{\texttt{assume}}
\UnaryInfC{$p$}
\RightLabel{\texttt{strip}}
\UnaryInfC{$p$}
\RightLabel{\texttt{negate}}
\UnaryInfC{$¬ p$}
\AxiomC{$p$}
\RightLabel{\texttt{canonicalize}}
\UnaryInfC{$p$}
\RightLabel{\texttt{simplify}}
\BinaryInfC{$⊥$}
\RightLabel{\texttt{canonicalize}}
\UnaryInfC{$⊥$}

\end{prooftree}
% \includegraphics[height=0.8\textheight]{figures/derivation.png}
\end{center}
\vfill
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=athena]{Athena}
  {\url{http://github.com/jonaprieto/athena}}

 \begin{tikzpicture}[overlay,remember picture]
 \tikzset{shift={(current page.north)}, xshift=0cm,yshift=0cm}
  \node[] at (0.92\textwidth, 1cm)
    {\includegraphics[width=0.15\textwidth]{figures/athena}};
\end{tikzpicture}

Is a Haskell program
that translates proofs given by Metis Prover
in \texttt{TSTP} format to Agda code.\\

\begin{itemize}
    \item Parsing of \texttt{TSTP} language
    \item Creation and analysis of \hyperlink{tstp-dag}{\textbf{DAG}} derivations
    \item Analysis of inference rules used in the \texttt{TSTP} derivation
    \item Agda code generation
\begin{table}[!ht]
\begin{center}
\scalebox{0.8}{
{\renewcommand{\arraystretch}{1.3}%
\begin{tabular}{|cl| }
\hline
\textbf{Library} & \textbf{Purpose} \\
\hline
\texttt{\hyperlink{agda-prop}{Agda-Prop}}& axioms and theorems of Classical Propositional Logic\\
\texttt{\hyperlink{agda-metis}{Agda-Metis}} & versions of the inference rules used by Metis\\
\hline
\end{tabular}}
}
\end{center}
\label{tab:library}
\end{table}

\begin{frame}[fragile, label=agda-prop-2]{Agda-Prop Library}
  {\url{https://github.com/jonaprieto/agda-prop}}
\begin{itemize}
  \item Propositional Logic + PEM ($Γ ⊢ ϕ ∨ ¬ ϕ$)
  \item A data type for formulas
\begin{agda}
data Prop : Set where
  Var  : Fin n → Prop           -- Variables.
  $⊤$    : Prop                   -- Top (truth).
  $⊥$    : Prop                   -- Bottom (falsum).
  _$∧$_  : ($φ$ $ψ$ : Prop) → Prop    -- Conjunction.
  _$∨$_  : ($φ$ $ψ$ : Prop) → Prop    -- Disjunction.
  _$⇒$_  : ($φ$ $ψ$ : Prop) → Prop    -- Implication.
  _$⇔$_  : ($φ$ $ψ$ : Prop) → Prop    -- Biimplication.
  $¬$_   : ($φ$ : Prop) → Prop      -- Negation.
\end{agda}
\end{itemize}
\end{frame}

\begin{frame}[fragile, label=agda-prop]{Agda-Prop Library}
  {\url{https://github.com/jonaprieto/agda-prop}}
\begin{itemize}
  \item A data type for theorems
\begin{agda}
data _$⊢$_ : ($Γ$ : Ctxt)($φ$ : Prop) → Set
\end{agda}

\item Constructors
\begin{agda}
assume, axiom, weaken, $⊤$-intro, $⊥$-elim, $¬$-intro,
$¬$-elim, $∧$-intro, $∧$-proj$₁$, $∧$-proj$₂$, $∨$-intro$₁$,
$∨$-intro$₂$, $∨$-elim, $⇒$-intro, $⇒$-elim, $⇔$-intro,
$⇔$-elim$₁$, $⇔$-elim$₂$
\end{agda}
  \item Natural deduction proofs for more than 71 theorems
\begin{agda}
$⇔$-equiv, $⇔$-assoc, $⇔$-comm, $⇒-⇔-¬∨$, $⇔-¬$-to-$¬$,
$¬⇔$-to-$¬$, $¬¬$-equiv, $⇒⇒-⇔-∧⇒$, $⇔$-trans, $∧$-assoc,
$∧$-comm, $∧$-dist, $¬∧$-to-$¬∨¬$, $¬∨¬$-to-$¬∧$, $¬∨¬-⇔-¬∧$,
subst$⊢∧₁$, subst$⊢∧₂$, $∨$-assoc, $∨$-comm, $∨$-dist,
$∨$-equiv, $¬∨$-to-$¬∧¬$, $¬∧¬$-to-$¬∨$, $∨$-dmorgan,
$¬¬∨¬¬$-to-$∨$, cnf, nnf, dnf, RAA, ...
\end{agda}
\end{itemize}
\end{frame}

\begin{frame}[fragile, label=agda-metis]{Agda-Metis Library}
  {\url{https://github.com/jonaprieto/agda-metis}}
\vfill
\begin{table}[!ht]
\begin{center}
\scalebox{0.7}{
{\renewcommand{\arraystretch}{1.3}%
\begin{tabular}{|clc|}
\hline
\textbf{Rule} & \textbf{Purpose} &\textbf{Theorem}\\
\hline
\texttt{canonicalize} & transforms formulas to CNF, DNF or NNF &\texttt{atp-canonicalize}\\
\texttt{clausify} & performs clausification &\texttt{atp-clausify}\\
\hyperlink{atp-conjunct}{\texttt{conjunct}} & takes a formula from a conjunction &\hyperlink{atp-conjunct}{\texttt{atp-conjunct}}\\
\texttt{negate} & applies negation to the formula &\texttt{atp-negate}\\
\texttt{resolve} & applies theorems of resolution &\texttt{atp-resolve}\\
\texttt{simplify} & applies over a list of formula to simplify them &\texttt{atp-simplify}\\
\texttt{strip} & splits a formula into subgoals &\texttt{strip}\\[1ex]
\hline
\end{tabular}}
}
\end{center}
\label{tab:agda-metis-table}
\end{table}
\vfill
\end{frame}

\begin{frame}[fragile, label=atp-conjunct]{Agda-Metis: Conjunct Inference}
   {\url{https://github.com/jonaprieto/agda-metis}}
\begin{itemize}
\item Definition
  \begin{equation*}
  conjunct (ϕ₁ ∧  ϕ₂ ∧ ⋯ ∧ \mathbf{ϕᵢ} ∧ ⋯ ∧ ϕ\_n, \mathbf{ϕᵢ}) ⟶ \mathbf{ϕᵢ}
  \end{equation*}

\item Function\footnote{\label{note-conjunct}Excerpt from the \texttt{Agda-Metis} library available in  \texttt{ATP.Metis.Rules.Conjunct} module.}:
\begin{agda}
conjunct : Prop → Prop → Prop
conjunct $(φ \wedge ψ)$ $ω$ with $⌊$ eq $φ$ $ω$ $⌋$ | $⌊$ eq ψ $ω$ $⌋$
... | true  | _     = $φ$
... | false | true  = $ψ$
... | false | false = conjunct $φ$ $ω$
conjunct $φ$ $ω$       = $φ$
\end{agda}
\item Theorem\textsuperscript{\ref{note-conjunct}}
\begin{agda}
atp-conjunct
  : ∀ $\{Γ\}$ $\{φ\}$
  → $(ω : Prop)$
  → $Γ \vdash φ$
  → $Γ \vdash$ conjunct $φ$ $ω$
\end{agda}
\end{itemize}
\end{frame}

\begin{frame}[fragile, label=atp-conjunct-proof]{A proof of \texttt{atp-conjunct} theorem}
\vfill
\begin{agda}
atp-conjunct
  : $∀ \{Γ\} \{φ\}$
  → ($ω$ : Prop)
  → $Γ ⊢ φ$
  → $Γ ⊢$ conjunct $φ\ ω$

atp-conjunct $\{Γ\}$ $\{φ ∧ ψ\}$ $ω$ $Γ⊢φ$ with $⌊$ eq $φ$ $ω$ $⌋$ | $⌊$ eq $ψ$ $ω$ $⌋$
... | true  | _     = $∧$-proj$₁$ $Γ⊢φ$
... | false | true  = $∧$-proj$₂$ $Γ⊢φ$
... | false | false =
    atp-conjunct $\{Γ = Γ\}$ $\{φ = φ\}$ $ω$ ($∧$-proj$₁$ $Γ⊢φ$)
atp-conjunct {_} {Var x} _ = id
atp-conjunct {_} {$⊤$} _     = id
atp-conjunct {_} {$⊥$} _     = id
atp-conjunct {_} {$φ ∨ φ₁$} _ = id
atp-conjunct {_} {$φ ⇒ φ₁$} _ = id
atp-conjunct {_} {$φ ⇔ φ₁$} _ = id
atp-conjunct {_} {$¬ φ$} _    = id
\end{agda}
\vfill

\end{frame}

@@

\begin{itemize}
  \item The problem is $p ∧ q ⊢ q ∧ p$
  \item In TPTP format
\begin{tptp}
fof(a, axiom, p & q).
fof(goal, conjecture, q & p).
\end{tptp}
  \item A natural deduction proof
\begin{columns}
\begin{column}{0.8\textwidth}
 \begin{prooftree}
  \AxiomC{$ϕ ∧ ψ$}
  \RightLabel{$∧$-\texttt{proj}$₁$}
  \UnaryInfC{$ϕ$}

@@

\begin{frame}[fragile, allowframebreaks]{TSTP derivation}
\begin{tptp}
fof(a, axiom, p & q).
fof(goal, conjecture, q & p).
fof(subgoal_0, plain, q,
    inference(strip, [], [goal])).
fof(subgoal_1, plain, q => p,
    inference(strip, [], [goal])).
fof(negate_0_0, plain, ~ q,
inference(negate, [], [subgoal_0])).
fof(normalize_0_0, plain, (~ q),
    inference(canonicalize, [], [negate_0_0])).
fof(normalize_0_1, plain, p & q,
    inference(canonicalize, [], [a])).
fof(normalize_0_2, plain, q,
    inference(conjunct, [], [normalize_0_1])).
fof(normalize_0_3, plain, $false,
    inference(simplify, [],
        [normalize_0_0, normalize_0_2])).
cnf(refute_0_0, plain, $false,
    inference(canonicalize, [], [normalize_0_3])).
fof(negate_1_0, plain, ~ (q => p),
    inference(negate, [], [subgoal_1])).
fof(normalize_1_0, plain, ~ p & q,
    inference(canonicalize, [], [negate_1_0])).
fof(normalize_1_1, plain, p & q,
     inference(canonicalize, [], [a])).
fof(normalize_1_2, plain, p,
    inference(conjunct, [], [normalize_1_1])).
fof(normalize_1_3, plain, q,
    inference(conjunct, [], [normalize_1_1])).
fof(normalize_1_4, plain, $false,
    inference(simplify, [],
      [normalize_1_0, normalize_1_2, normalize_1_3])).
cnf(refute_1_0, plain, ($false),
    inference(canonicalize, [], [normalize_1_4])).
\end{tptp}
\end{frame}

\begin{frame}[fragile, label=verified-example]{$p ∧ q ⊢ q ∧ p$}{Reconstructed proof}
\vfill
Definitions
\begin{agda}
p : Prop
p = Var (# 0)
q : Prop
q = Var (# 1)
-- Axiom.
a : Prop
a = (p $∧$ q)
-- Premise.
$Γ$ : Ctxt
$Γ$ = [ a ]
-- Conjecture.
goal : Prop
goal = (q $∧$ p)
-- Subgoals.
subgoal$₀$ : Prop
subgoal$₀$ = q
subgoal$₁$ : Prop
subgoal$₁$ = (q $⇒$ p)
\end{agda}
\end{frame}

\begin{frame}[fragile, label=verified-example-2]{$p ∧ q ⊢ q ∧ p$}{Reconstructed proof}
\vfill
\begin{agda}
-- Axiom.
a : Prop
a = (p $∧$ q)
-- Subgoal.
subgoal$₀$ : Prop
subgoal$₀$ = q

proof$₀$ : $Γ ⊢$ subgoal$₀$
proof$₀$ =
  (RAA
    (atp-canonicalize
      (atp-simplify
        (atp-canonicalize
          (atp-strip
            (assume $\{Γ = Γ\}$ (atp-negate subgoal$₀$))))
        (atp-conjunct (q)
          (atp-canonicalize
            (weaken (atp-negate subgoal$₀$)
              (assume $\{Γ = ∅\}$ a)))))))
\end{agda}
\vfill
\end{frame}

\begin{frame}[fragile, label=verified-example-3]{$p ∧ q ⊢ q ∧ p$}{Reconstructed proof}
\vfill
\begin{agda}
subgoal$₁$ : Prop
subgoal$₁$ = (q $⇒$ p)

proof$₁$ : $Γ ⊢$ subgoal$₁$
proof$₁$ =
  (RAA
    (atp-canonicalize
      (atp-simplify
        (atp-conjunct (q)
          (atp-canonicalize
            (weaken (atp-negate subgoal$₁$)
              (assume $\{Γ = ∅\}$ a))))
        (atp-simplify
          (atp-canonicalize
            (atp-strip
              (assume $\{Γ = Γ\}$ (atp-negate subgoal$₁$))))
          (atp-conjunct (p)
            (atp-canonicalize
              (weaken (atp-negate subgoal$₁$)
                (assume $\{Γ = ∅\}$ a))))))))
\end{agda}
\vfill
\end{frame}

\begin{frame}[fragile, label=verified-example-4]{$p ∧ q ⊢ q ∧ p$}{Reconstructed proof}
\begin{agda}
proof : $Γ ⊢$ goal
proof =
  $⇒$-elim
    atp-splitGoal
    ($∧$-intro proof$₀$ proof$₁$)
\end{agda}
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=failure-example]{Bug in the Printing of the Proof}
  {Metis' Issue: \url{https://github.com/gilith/metis/issues/2}}
% {}
% \begin{tptp}
% $ cat problem.tptp
% fof(goal, conjecture,
%   ((p <=> q) <=> r) <=> (p <=> (q <=> r))).
% \end{tptp}
\begin{tptp}
$ metis --version
metis 2.3 (release 20161108)
$ metis --show proof problem.tptp
...
fof(normalize_2_0, plain,
  (~ p & (~ q <=> ~ r) & (~ p <=> (~ q <=> ~ r))),
  inference(canonicalize, [], [negate_2_0])).
fof(normalize_2_1, plain, ~ p <=> (~ q <=> ~ r),
    inference(conjunct, [], [normalize_2_0])).
fof(normalize_2_2, plain, ~ q <=> ~ r,
    inference(conjunct, [], [normalize_2_0])).
fof(normalize_2_3, plain, ~ p,
  inference(conjunct, [], [normalize_2_0])).
fof(normalize_2_4, plain, $false,
    inference(simplify, [],
      [normalize_2_1, normalize_2_2, normalize_2_3])).
...
\end{tptp}
\end{frame}

\begin{frame}[fragile, label=hammer]{Related Work}
\begin{block}{SledgeHammer} \citep{Paulson2007}
\begin{itemize}
\item Isabelle/HOL mature tool
\item Metis ported within Isabelle
\item Reconstruct proofs of well-known ATPs: \texttt{EProver}, \texttt{Vampire},
among others using SystemOnTPTP server
\end{itemize}
\end{block}

\begin{block}{Integrating Waldmeister into Agda} \citep{Foster2011}
\begin{itemize}
  \item Framework for a integration between Agda and ATPs
  \begin{itemize}
    \item Equational Logic
    \item Reflection Layers
  \end{itemize}
  \item Source code is not available\footnote{\url{http://simon-foster.staff.shef.ac.uk/agdaatp}}
  \end{itemize}
\end{block}
\end{frame}

@@

@@

\only<4->{
  \node[below= 0.4cm of eagda] (tptp)
    {\includegraphics[scale=0.8]{figures/tptp}}};

@@

\only<7->{ \node[left= 2.75cm of atp, rectangle, text=white,fill=black, align=left, font=\tiny] (proved)
{
\texttt{\$ agda Or.agda}\\
\texttt{\$ apia Or.agda --atp=online-metis}\\
\texttt{Metis---2.3 proved the conjecture}
  }
};

@@

\begin{frame}[label=pending-work]{Pending Work}
\begin{itemize}
\item There are missing cases with the \texttt{simplify} inference
\item Is not clear, how  \texttt{canonicalize} inference choose what
normal form use to transform the formulas
\item When Metis prove a goal, it first splits that goal in a list of
subgoal, then we are working on verification of such splitting.
\end{itemize}
\end{frame}

\begin{frame}[label=future-work]{Future Work}
\begin{itemize}
  \item Add shallow embedding in order to work with \texttt{Apia}
  \item Support First-Order Logic with Equality
  \item Support another prover like EProver or Vampire
\end{itemize}
\end{frame}
